Introducing ICBe: Very High Recall and Precision Event Extraction from Narratives about International Crises

### 2023-06-03 replication_paper/psrm_draft/ICBe_manuscript.Rmd
ontology innovation para added. intro header deleted

# Introduction

@@

Our contribution is using human coders and natural language processing to develop (1) a large, flexible ontology of international affairs and (2) a fine-grained and structured event dataset of international crises from 1918-2017 developed by applying the ontology to an unusually high-quality corpus of historical narratives of international crises [@brecherInternationalStudiesTwentieth1999; @wilkenfeldInterstateCrisesViolence2000; @brecherInternationalCrisisBehavior2021]. We then develop several methods for objectively gauging how well these event codings reconstruct the information contained in the original crisis narrative. We conclude by benchmarking our event codings against several current state of the art event data collection efforts. We find that existing systems produce sequences of events that do not contain enough information to reconstruct the underlying historical episode. The underlying fine-grained variation in international affairs is unrecognizable through the lens of current quantification efforts. In focusing this initial effort on international crises as a proof of concept sample, we demonstrate our ontology and method's potential to improve upon existing empirical identifications of patterns of international interactions.

@@

Table 1 demonstrates that with the exception of large-scale CAMEO dictionary-based systems, our ontology improves upon the the existing state of the art quantitative datasets that ignore important information content about international interactions.[^5] In particular,

@@

-   separating chess pieces from players (military vs civilian)

-   addition of domains

>>>>

@@

We advance existing efforts to identify and structure regularized events and actor in international politics by using human coders and natural language processing to develop (1) a large, flexible ontology of international affairs and (2) a fine-grained and structured event dataset of international crises from 1918-2017 developed by applying our ontology to an unusually high-quality corpus of historical narratives of international crises [@brecherInternationalStudiesTwentieth1999; @wilkenfeldInterstateCrisesViolence2000; @brecherInternationalCrisisBehavior2021]. We then develop several methods for objectively gauging how well these event codings reconstruct the information contained in the original crisis narrative. We conclude by benchmarking our event codings against several current state of the art event data collection efforts. We find that existing systems produce sequences of events that do not contain enough information to reconstruct the underlying historical episode. The underlying fine-grained variation in international affairs is unrecognizable through the lens of current quantification efforts. In focusing this initial effort on international crises as a proof of concept sample, we demonstrate our ontology and method's potential to improve upon existing empirical identifications of patterns of international interactions.

@@

Table 1 demonstrates that with the exception of large-scale CAMEO dictionary-based systems, our ontology improves upon the the existing state of the art quantitative datasets that ignore important information content about international interactions.[^5] In particular, we highlight two innovations in our ontology. First, we separate the chess pieces from the chess players in not only identifying what states were involved in each crisis, but distinguishing between different actors within a state. By virtue of our ontology coding military versus civilian actors and national leaders versus bureaucrats, our data can be used to explore important questions concerning topics like civilian military relations as well as evolution in what actors are engaged in crises - a topic of increasing interest as states engage in "gray zone conflict" by employing "chess pieces" like the coast guard instead of military navy. Second, we add information about the domains in which actors behave - whether in land, air, sea, space, or cyber - since they differ in their technology, tactics, geography, and purpose [@gartzkeCrossDomainDeterrenceStrategy2019]. Doing so allows researchers to identify and explain patterns in escalation conditional on the military means states use in conflict. Recent concerns about cross-domain conflict and the effect of new domains of conflict like space and cyber have made this an endeavor of increased interest to practitioners [@gannonOneIfLand2022].

@@
